Adversarial Deep Neural Networks Effectively Remove Nonlinear Batch Effects from Gene-Expression Data

### 2019-07-04 thesis.tex
Fix error to make latex compile

Recent progress has been made in using artificial neural networks to correct batch effects in expression data \cite{shaham_removal_2017,shaham_batch_2018,upadhyay_removal_2019}.
However, these methods are designed for a limited range of scenarios:
1) when the input data contain only two batches, 
2) the batches are sufficiently large (> 100 samples in our testing), and 
3) the batches are balanced.[@For #2, does this mean the total number of samples? Or the number of samples per batch? Please clarify the wording.@] 
These requirements rarely hold in applied research; for example, the bladderbatch dataset---which is used to test batch-adjustment techniques\cite{leek_bladderbatch_2017,leek_sva_2017}---has 5 batches, 57 samples, and 4 to 19 samples per batch.
Additionally, existing methods for batch correction have not been tested to verify whether nonlinear patterns remain in expression data after batch correction.
To address these limitations, we implemented a deep neural network that [@describe our method briefly, what is different about our method, and why we think our method can be applied when there are >2 batches, small datasets, and unbalanced batches.@] 
Furthermore, we tested our method using machine-learning classification algorithms to quantify the extent to which nonlinear batch effects remained after adjustment.

>>>>

Recent progress has been made in using artificial neural networks to correct batch effects in expression data \cite{shaham_removal_2017,shaham_batch_2018,upadhyay_removal_2019}.
However, these methods are designed for a limited range of scenarios:
1) when the input data contain only two batches,
2) the batches are sufficiently large (> 100 samples in our testing), and
3) the batches are balanced.[@For \#2, does this mean the total number of samples? Or the number of samples per batch? Please clarify the wording.@]
These requirements rarely hold in applied research; for example, the bladderbatch dataset---which is used to test batch-adjustment techniques\cite{leek_bladderbatch_2017,leek_sva_2017}---has 5 batches, 57 samples, and 4 to 19 samples per batch.
Additionally, existing methods for batch correction have not been tested to verify whether nonlinear patterns remain in expression data after batch correction.
To address these limitations, we implemented a deep neural network that [@describe our method briefly, what is different about our method, and why we think our method can be applied when there are >2 batches, small datasets, and unbalanced batches.@]
Furthermore, we tested our method using machine-learning classification algorithms to quantify the extent to which nonlinear batch effects remained after adjustment.
